---
title: "Coursera FARS package"
author: "Nobuhiro Kikuchi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Coursera FARS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, echo = FALSE, include = FALSE}
library(fars)
library(dplyr)
library(maps)
```


This is a vignette for the FARS pacage functions. FARS pacakage offers a set of functions for analysis of data from FARS - Fatality Analysis Reporting System. FARS is a nationwide census providing NHTSA, Congress and the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. [[1](http://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS))]

## Package functions

Note:
the data file needs to be present in the working directory. 

There are three exported functions available to users:

- `fars_read()`
- `fars_summarize_years()`
- `fars_map_state()`

 Follwing examples explain each function. We will use data from FARS, years 2013-2015, that can be found in the package directory under `\extdata` folder.

## Read FARS data files

The function `fars_read()` provides a way how to read data from a file path provided as an argument. It returns a `tibble` with loaded data:

```{r fars_read_example}
filename <- system.file("extdata/accident_2013.csv.bz2", package = "fars")
fars_read(filename)
```

## Summarize number of accidents


`fars_summarize_years()` takes a vector or list of years (numeric/integer values) as an argument. It returns a pivot table with number of accidents per year and month, if the corresponding files with data are available. 

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "fars"))
fars_summarize_years(2013:2015)
```

## Plot accident locations

`fars_map_state()` takes a state index and year as arguments. It returns a plot of the state with accident locations. The state indices need to correspond available indeces from the data set. 


```{r fars_map_state_example}
setwd(system.file("extdata", package = "fars"))
fars_map_state(45, 2015)
```


